One page description

### 2016-12-01 publications/Journal/1-introduction.tex
Summary of changes: - add proposed reply to reviewers - update abstract, introduction to mention PDEVS explicitly as a protocol compl$ - detail in a subsection implementation of PDEVS - detail our usage of parameter q Complete todo from previous review.
First changes after review
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

This paper introduces \textit{DEVS-Ex-Machina}\footnote{\url{https://bitbucket.org/bcardoen/devs-ex-machina}} (``\textit{dxex}''): our simulation tool~\cite{dxex} which offers multiple synchronization protocols: no synchronization (sequential execution), conservative synchronization, or optimistic synchronization.
The selected synchronization protocol is transparent to the simulated model: users only determine the protocol to use.
Users who simulate a wide variety of models, with different ideal synchronization protocols, can run the exact same model using the exact same tool, but with different synchronization protocols.
As model behaviour, and thus the ideal synchronization protocol, might change during simulation, runtime switching of synchronization protocols is also supported.
This runtime switching can be based on performance metrics, which are logged during simulation.
Information is made available to a separate component, where a choice can be made about which synchronization protocol to use. % It's manual now, can be extended to full automatic.
Additionally, we investigate how model allocation influences the performance of our synchronization protocols.
To this end, we have included an allocation component in our simulation kernel.

>>>>

This paper introduces \textit{DEVS-Ex-Machina}\footnote{\url{https://bitbucket.org/bcardoen/devs-ex-machina}} (``\textit{dxex}''): our simulation tool~\cite{dxex} which offers multiple synchronization protocols: no synchronization (single core execution), conservative synchronization, or optimistic synchronization. Each inter core synchronization protocol can be complemented with the parallel execution of concurrent events within a simulation core. We refer to this intra core protocol as pSim for the remainder of this paper. \\
This results in the following list of available parallelization approaches:\\
\begin{itemize}
\item No inter core synchronization (single core), no intra core synchronization : S
\item Conservative inter core synchronization, no intra core synchronization : C
\item Optimistic inter core synchronization, no intra core synchronization : O
\item No inter core synchronization (single core), pSim intra core synchronization : SP
\item Conservative inter core synchronization, pSim intra core synchronization : CP
\item Optimistic inter core synchronization, pSim intra core synchronization : OP
\end{itemize}
The selected synchronization protocol is transparent to the simulated model: users only determine the protocol to use.
Users who simulate a wide variety of models, with different ideal synchronization protocols, can run the exact same model using the exact same tool, but with different synchronization protocols.
As model behaviour, and thus the ideal synchronization protocol, might change during simulation, runtime switching of synchronization protocols is also supported.
This runtime switching can be based on performance metrics, which are logged during simulation.
Information is made available to a separate component, where a choice can be made about which synchronization protocol to use. % It's manual now, can be extended to full automatic.
Additionally, we investigate how model allocation influences the performance of our synchronization protocols.
To this end, we have included an allocation component in our simulation kernel.

### 2016-12-01 publications/Journal/3-features.tex
First changes after review
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

\subsubsection{Sequential}
Our sequential simulation algorithm is very similar to the one found in \textit{PythonPDEVS}, including many optimizations.
Minor modifications were made, though, such that it can be overloaded by different synchronization protocol implementations.
This way, the \textsf{DEVS} simulation algorithm is implemented once, but parts can be overridden as needed.
In theory, more synchronization protocols (\textit{e.g.}, other algorithms for conservative synchronization) can be added without altering our design.

An overview of \textit{dxex}'s design is given in Figure~\ref{fig:class_diagram}.
It shows that there is a simulation \texttt{Core}, which simulates the \texttt{AtomicModel}s connected to it.
The superclass \texttt{Core} represents the sequential simulation kernel.
Subclasses define specific variants, such as \texttt{ConservativeCore} (conservative synchronization), \texttt{OptimisticCore} (optimistic synchronization), and \texttt{DynamicCore} (\textsf{Dynamic Structure DEVS}).

@@

% protocol or inherent parallelism?
\subsubsection{PDEVS}
In our tool each synchronization kernel is capable of executing its transitions in parallel. A configurable number of threads is allocated to optimally divide the processing load of concurrent transitions. The parallel execution of transitions introduces some overhead in thread pooling and locking, and disallows some optimizations in optimally rescheduling models. The concurrency of events and the combined computational load of the transitions has to outweigh this overhead to obtain a significant speedup, as we will demonstrate in Section~\textsc{\nameref{sec:4-performance}}.

>>>>

\subsubsection{Single Core}
Our single core simulation algorithm is very similar to the one found in \textit{PythonPDEVS}, including many optimizations.
Minor modifications were made, though, such that it can be overloaded by different synchronization protocol implementations.
This way, the \textsf{DEVS} simulation algorithm is implemented once, but parts can be overridden as needed.
In theory, more synchronization protocols (\textit{e.g.}, other algorithms for conservative synchronization) can be added without altering our design.

An overview of \textit{dxex}'s design is given in Figure~\ref{fig:class_diagram}.
It shows that there is a simulation \texttt{Core}, which simulates the \texttt{AtomicModel}s connected to it.
The superclass \texttt{Core} represents the single core simulation kernel.
Subclasses define specific variants, such as \texttt{ConservativeCore} (conservative synchronization), \texttt{OptimisticCore} (optimistic synchronization), and \texttt{DynamicCore} (\textsf{Dynamic Structure DEVS}).

@@

% protocol or inherent parallelism?
\subsubsection{pSim}
In our tool each simulation core, regardless of inter core synchronization, is capable of executing concurrent transitions in parallel. A configurable number of threads is allocated to optimally divide the processing load of concurrent transitions. The parallel execution of transitions introduces some overhead in thread pooling and locking, and disallows some optimizations in optimally rescheduling models. The concurrency of events and the combined computational load of the transitions has to outweigh this overhead to obtain a significant speedup, as we will demonstrate in Section~\textsc{\nameref{sec:4-performance}}.

### 2016-12-01 publications/Journal/4-performance.tex
First changes after review
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

\subsection{Comparison to the abstract simulator}
The abstract simulator of \textsf{Parallel DEVS} included its own notion of parallelism: transition functions that happen at the same point in simulated time can be trivially parallelized.
This parallelization is trivial, in the sense that there is no need for synchronization protocols.
The parallelization of the transition functions (from now on called the ``inherent parallelism'') is, at least in \textit{dxex}, independent of the use of a synchronization protocol.
That is, both can be used simultaneously.
Following the theoretical analysis published in~\cite{amdahlpdevs}, a comparison between this inherent parallelism and the use of synchronization protocols is warranted.

\paragraph{Model}
We have opted to use the \textit{Queue} model for this comparison, as it allows for many interesting configurations.
%TODO what is meant with this allocation of threads and synchronized kernels?
% 4 Kernels running on a dedicated thread, with per kernel 4 threads allocated to use for PDEVS.
The effect of combining 4 synchronized kernels, with each 4 threads for PDEVS, is contrasted with 16 threads for PDEVS and 16 kernels for conservative and optimistic.
The first scenario runs each kernel on a dedicated thread synchronized with the other 3 kernels. Each kernel can use up to 4 threads for PDEVS. In the second scenario PDEVS can use up to 16 threads while running in a sequential kernel. Finally 16 synchronized kernels are used with one thread per kernel.
This allows us to observe which is more efficient in obtaining a speedup.
We simulate a computational load by an optional call to sleep in the transition functions set at 5 milliseconds.
The model is configured with depth 4 and width 300 if the transition function has no load, and width 50 if the computational load is active.
In our configuration, an imminent model generates output, resulting in the receiving model becoming imminent.
With respect to the theoretical analysis ~\cite{amdahlpdevs} this gives a parameter value $q = 1$ for a fixed timeadvance and $q=0.5$ for a random timeadvance. Q is the probability that a receiving model becomes immediately imminent. 
Without a fixed timeadvance the receiving model becomes immediately imminent. 
With random scheduling the receiving model becomes imminent with a scheduled time in the current timeslot or the next, chosen with a uniform distribution, therefore $q = \frac{1}{2}$.
 The benchmarks are run sufficiently long enough to guarantee that the frequency of internal and external transitions is equal between random and non-random benchmarks. The key difference is that although the event frequency is the same, their concurrency is not.
Models are equally distributed over the kernels and between threads in PDEVS.
The coefficient of variation of our results is less than 1\%.
The communication overhead is hard to estimate, but given our coefficient of variation, we can expect this overhead to be constant.

>>>>

\subsection{Comparison to the abstract simulator}
%reference pSim
The abstract simulator of \textsf{Parallel DEVS} included its own notion of parallelism: transition functions that happen at the same point in simulated time can be trivially parallelized. In this work we refer to this approach as pSim.
This parallelization is trivial, in the sense that there is no need for synchronization protocols.
The parallelization of the transition functions (from now on called the ``inherent parallelism'') is, at least in \textit{dxex}, independent of the use of a synchronization protocol.
That is, both can be used simultaneously.
Following the theoretical analysis published in~\cite{amdahlpdevs}, a comparison between this inherent parallelism and the use of synchronization protocols is warranted.

\paragraph{Model}
We have opted to use the \textit{Queue} model for this comparison, as it allows for many interesting configurations.
%TODO what is meant with this allocation of threads and synchronized kernels?
% 4 Kernels running on a dedicated thread, with per kernel 4 threads allocated to use for PDEVS.
The effect of combining 4 synchronized kernels, with each 4 threads for PDEVS, is contrasted with 16 threads for PDEVS and 16 kernels for conservative and optimistic.
The first scenario runs each kernel on a dedicated thread synchronized with the other 3 kernels. Each kernel can use up to 4 threads for PDEVS. In the second scenario PDEVS can use up to 16 threads while running in a sequential kernel. Finally 16 synchronized kernels are used with one thread per kernel.
This allows us to observe which is more efficient in obtaining a speedup.
We simulate a computational load by an optional call to sleep in the transition functions set at 5 milliseconds.
The model is configured with depth 4 and width 300 if the transition function has no load, and width 50 if the computational load is active.
In our configuration, an imminent model generates output, resulting in the receiving model becoming imminent.
The probability that an internal transition in a model generates output and is connected to a receiving model here is 1. With respect to the theoretical analysis ~\cite{amdahlpdevs} this gives a parameter value $q = 1$. 
The probability that a model becomes imminent (i.e. is scheduled to execute an internal transition at a given point in time) in this model is equal to 1 when a fixed timeadvance is used,  %todo p calculation
$\frac{1}{timestep}$ when a random time advance is used, this corresponds with $p = 1$ and $p = \frac{1}{100}$ respectively.

The benchmarks are run sufficiently long enough to guarantee that the frequency of internal and external transitions is equal between random and non-random benchmarks. In our model each model will be executing internal and external transitions, creating an ideal use case to evaluate the speedup inherent parallelism can obtain.
The key difference is that although the event frequency is the same, their concurrency is not. 
Models are equally distributed over the kernels and between threads in PDEVS.
The coefficient of variation of our results is less than 1\%.
The communication overhead is hard to estimate, but given our coefficient of variation, we can expect this overhead to be constant.
